---
title: "Hospitals"
author: "Andre Williams"
date: '2022-06-13'
output: html_document
---

# The  hospitals data set.
Reading the Data set in.

```{r}
library(tidyverse)
hospitals <- read.csv("http://www.stat.cmu.edu/cmsac/sure/2022/materials/data/health/eda_projects/hospitals.csv")
head(hospitals)
```


# Hypotheses 1: Government hospitals have the worse overall rating compared to other institutions.

1) Grouping the data set by the facility type.
2) Averaging the ratings of each facility type.
3) Displaying the new table with the facility type and average overall rating.

```{r}
hospital_ratings <- hospitals %>%
  group_by(Facility.Type) %>%
  summarize(Overall.Ratings = mean(Rating.Overall))

hospital_ratings

```

4) Plotting the average hospital ratings by facility.

```{r}
hospital_ratings %>%
  ggplot(aes(x = Facility.Type, y = Overall.Ratings, fill = Facility.Type)) +
  geom_col(width = 0.5) +
  theme_bw()


```


Our first hypotheses was deemed correct, as government hospitals had an overall rating of ~1.5 stars.




# Hypotheses #2: States in the Northeast tend to have a high cost of procedures.

1)Check the population for each state

```{r}
census <- read.csv("https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/state/detail/SCPRC-EST2019-18+POP-RES.csv")
head(census)
```

```{r}
population0 <- as_tibble(census) %>%
  select(NAME, POPESTIMATE2019)

colnames(population0)[1] <- 'state'
colnames(population0)[2] <- 'population'

population0 <- population0[-c(1, 53),]
```

```{r}
library(usmap)
library(ggplot2)

plot_usmap(data = population0, values = "population", color = "red") + 
  scale_fill_continuous(low = "yellow", high = "darkred", name = "population", label = scales::comma) + 
  theme(legend.position = "right")

```

2) Group the hospitals by their states.

```{r}
hospital_cost <- hospitals %>%
  group_by(Facility.State)
```

3) Calculate the average cost of each procedures per state.

```{r}
hospital_cost <- hospital_cost %>%
  summarise(Avg_HCost = mean(Procedure.Heart.Attack.Cost),
            Avg_HFCost = mean(Procedure.Heart.Failure.Cost),
            Avg_PCost = mean(Procedure.Pneumonia.Cost),
            Avg_HKCost = mean(Procedure.Hip.Knee.Cost))

colnames(hospital_cost)[1] <- 'state'
```

4) Combine the average of each procedure to obtain one average cost of all.

```{r}
hospital_avg_cost <- hospital_cost %>%
  group_by(state) %>%
  summarize(Avg_cost = mean(Avg_HCost + Avg_HFCost + Avg_PCost + Avg_HKCost)) %>%
  arrange(desc(Avg_cost))
  
hospital_avg_cost
```

5) Plot the map with the average cost of procedures for each state.

```{r}
library(usmap)
library(ggplot2)

plot_usmap(data = hospital_avg_cost, values = "Avg_cost", color = "red") + 
  scale_fill_continuous(low = "yellow", high = "darkred", name = "Avg Cost", label = scales::comma) + 
  theme(legend.position = "right")


```





With the exception of Florida, our hypotheses was correct, as New Jersey and the rest of northeast America have a higher cost of living. 




# Hypotheses 3: The average overall cost of all four procedures will be lowest in Church Facility Type hospitals. Cameron has the more complete code in his Rstudio. I just have some of the calculations below.

#Making the variable 'all_cost' in my 'ratings' dataset  

#ratings_nozero has all of the aprx. 2,500 zero values filtered out
#ratings has a new column for all_cost0 that keeps the zeroes in while also taking the average cost across all four procedures per hospital in the 'ratings' dataset

ratings_nozero <- hospitals %>%
  filter(`Procedure.Heart Attack.Cost` > 0, `Procedure.Heart Failure.Cost` > 0, Procedure.Pneumonia.Cost > 0, `Procedure.Hip Knee.Cost` > 0) %>%
  mutate(all_cost = (`Procedure.Heart Attack.Cost` + `Procedure.Heart Failure.Cost` + `Procedure.Pneumonia.Cost` + `Procedure.Hip Knee.Cost`) / 4)

ratings <- hospitals %>%
  mutate(all_cost0 = (`Procedure.Heart Attack.Cost` + `Procedure.Heart Failure.Cost` + `Procedure.Pneumonia.Cost` + `Procedure.Hip Knee.Cost`) / 4)

#Violin-Boxplot

#Violin-Boxplot of the mean of all procedure costs (including Zeroes) in our hospital dataset
hospitals %>%
  ggplot(aes(x = Facility.Type,
             y = all_cost0)) +
  geom_violin() +
  geom_boxplot(width = .2) +
  theme_bw()

#Violin-Boxplot of the mean of all procedure costs but with Zeroes filtered out
ratings_nozero %>%
  ggplot(aes(x = Facility.Type,
             y = all_cost)) +
  geom_violin(aes(fill = factor(Facility.Type))) +
  geom_boxplot(width = .2) +
  theme_bw()


#A Stacked Bar Plot showing the distribution of frequency of the 4 procedures (Supporting plot to the 2 Violin-Boxplots)
ratings_sub = hospitals[c(4,13,16,19,22)]
procedure_names = c("Empty","Heart Attack", "Heart Failure", "Pneumonia", "Hip Knee")
type = c()
procedure = c()
n = dim(hospitals)[1]
for (i in 1:n){
  for (j in 2:5){
    if (as.numeric(ratings_sub[i,j])>0){
      type = c(type, as.character(ratings_sub[i,1]))
      procedure = c(procedure, procedure_names[j])
    }
  }
}
type_procedure = data.frame("Type" = type, "Procedure" = procedure)
type_procedure %>% ggplot(aes(x = Type, fill = Procedure)) +
  geom_bar() +
  labs(title = " ",
       x = "Type", y = "Procedure")


----------------------------------------------------------------------------------------------------------------------------------------------------------

# Clustering Example: K-Means Clsutering

clean_ratings <- hospitals %>%
  filter(`Procedure.Heart Attack.Cost` > 0, `Procedure.Heart Failure.Cost` > 0)
clean_ratings

#Standardize the variables

clean_ratings <- clean_ratings %>%
  mutate(std_log_cost = as.numeric(scale(`Procedure.Heart Attack.Cost`, center = TRUE, scale = TRUE)),
         std_rating_over = as.numeric(scale(`Procedure.Heart Failure.Cost`, center = TRUE, scale = TRUE)))

#K-means Algorithm with 2 centroids (We can compare with the 3 centers clustering algorithm)
std_kmeans <- 
  kmeans(dplyr::select(clean_ratings, std_log_cost, std_rating_over),
         algorithm = "Lloyd", centers = 2, nstart = 1)
         
#We can show a side-by-side clustering plot with 3 centroids

std_kmeans2 <- 
  kmeans(dplyr::select(clean_ratings, std_log_cost, std_rating_over),
         algorithm = "Lloyd", centers = 3, nstart = 1)

#Clustering Alogrithm for Heart Attack Cost and Heart Failure Cost

clean_ratings %>%
  mutate(rating_clusters = 
           as.factor(std_kmeans$cluster)) %>%
  ggplot(aes(x = `Procedure.Heart Attack.Cost`, y = `Procedure.Heart Failure.Cost`,
             color = rating_clusters)) +
  geom_point() + 
  ggthemes::scale_color_colorblind() +
  theme_bw() +
  theme(legend.position = "bottom") +
  coord_fixed()
  
#Second clustering example

clean_ratings %>%
  mutate(rating_clusters = 
           as.factor(std_kmeans2$cluster)) %>%
  ggplot(aes(x = `Procedure.Heart Attack.Cost`, y = `Procedure.Heart Failure.Cost`,
             color = rating_clusters)) +
  geom_point() + 
  ggthemes::scale_color_colorblind() +
  theme_bw() +
  theme(legend.position = "bottom") +
  coord_fixed()
  
  
```{r}
library(plyr)
hospitals$Procedure.Heart.Attack.Quality <- revalue(hospitals$Procedure.Heart.Attack.Quality, c("Better" = 1, "Unknown" = -1, "Lower" = 2, "Worse" = 4, "Average" = 3))
hospitals <- hospitals %>%
  mutate_at(vars(Procedure.Heart.Attack.Quality), as.numeric)

```

```{r}
hospitals %>%
  select(Facility.State, Procedure.Heart.Attack.Quality)
```

```{r}
hospitals_costperquality <- hospitals %>%
  mutate(hospital_cpq = Procedure.Heart.Attack.Cost / Procedure.Heart.Attack.Quality) %>%
  select(Facility.State, hospital_cpq)

hospitals_costperquality

```
